SETI in the Optical Spatial-Temporal Domain

### 2019-04-09 paper/ms.tex
add a paragraph monday

\documentclass[modern]{aastex62}
\bibliographystyle{aasjournal}

@@

Traditional SETI work also requires time-consuming observations, often through dedicated monitoring of nearby stars by radio telescopes. Such observing campaigns are expensive and difficult to obtain given the resource-limited nature of telescope allocation. Recent progress for a systematic ``technosignature'' search has been made by Breakthrough Listen \citep{worden2017,isaacson2017}, primarily at radio wavelengths \citep[e.g.][]{price2018}.

@@

Surveys like the currently running Zwicky Transient Facility \citep[ZTF][]{bellm2014} and the upcoming Large Synoptic Survey Telescope \citep[LSST][]{lsst} will also be transformative in enabling both reproducible and cost-efficient SETI. By developing search strategies that utilize public survey databases and real-time ``alert streams'', both of which are being developed to facilitate a wide range of science goals from these facilities, optical survey SETI can be conducted automatically. As new search algorithms are developed, these databases can be quickly re-analyzed, providing a much-needed level of reproducibility and transparency that will help reduce the ``giggle factor'' stigma for SETI \citep{wright2018b}.

In this paper I explore the potential of wide-field time domain surveys for conducting a new generation of technosignature searches. 
In \S\ref{sec:method} I describe the general principle behind technosignature searches in time domain surveys, including spatial over-densities of light curve signals and coordinated signals between multiple sources, for example.
In \S\ref{sec:haystack} I introduce several relevant surveys, and use the ``Cosmic Haystack'' search volume metric developed by \citet{wright2018c} to quantitatively compare these surveys to current radio SETI approaches. 
The data-driven spatial-temporal domain explored by these surveys necessitates new SETI approaches be developed, and in \S\ref{sec:transit} I outline an example signal where transiting planets around many stars could act as a spatially distributed beacon network.
Finally in \S\ref{sec:discussion} I discuss ideas for other search approaches, and the advantages in cost and reproducibility of conducting SETI with current and upcoming optical time domain surveys.

%Many have noted these new surveys can be used for SETI, such as looking for unusual transits (REFS), and they have the potential to produce viable, repeatable SETI searches from surveys if the techniques can be formed into survey or database queries and therefore run on nearly any survey, such as searching for ``impossible'' patterns or sequences in light curves (e.g. Davenport et al in prep)

@@

%In this work we propose a new type of beacon system, which relies on ET transit signals from multiple stars to collectively ``point'' towards an ET civilization. This system has the advantage of being potentially detectable in the near future via exoplanet searches.

%%%%
\section{Optical Survey SETI}
\label{sec:method}
METHOD: what is the big idea, in detail

include mantras of "worth doing", "do it almost for free", "portable, do it everywhere", and "outliers are nature papers"

>>>>

\documentclass[twocolumn]{aastex62}
\bibliographystyle{aasjournal}

@@

Traditional SETI work also requires time-consuming observations, often through dedicated monitoring of nearby stars by radio telescopes. Such observing campaigns are expensive and difficult to obtain given the competitive nature of telescope allocation. Recent progress for systematic ``technosignature'' searchs has been made by Breakthrough Listen \citep{worden2017,isaacson2017}, primarily at radio wavelengths \citep[e.g.][]{price2018}.

@@

Surveys like the currently running Zwicky Transient Facility \citep[ZTF][]{bellm2014} and the upcoming Large Synoptic Survey Telescope \citep[LSST][]{lsst} will be transformative in enabling both reproducible and cost-efficient SETI. By developing search strategies that utilize public survey databases and real-time ``alert streams'', both of which are being developed to facilitate a wide range of science goals from these facilities, optical survey SETI can be conducted automatically. As new search algorithms are developed, these databases can be quickly re-analyzed, providing a much-needed level of reproducibility and transparency that will help reduce the ``giggle factor'' stigma for SETI \citep{wright2018b}.

In this paper I explore the potential for wide-field time domain surveys in conducting a new generation of technosignature searches. 
In \S\ref{sec:method} I describe the general principles behind technosignature searches in time domain surveys, including spatial over-densities of light curve signals and coordinated signals between multiple sources, for example.
In \S\ref{sec:haystack} I introduce several relevant surveys, and use the ``Cosmic Haystack'' search volume metric developed by \citet{wright2018c} to quantitatively compare these surveys to current radio SETI approaches. 
The data-driven spatial-temporal domain explored by these surveys necessitates new SETI approaches be developed, and in \S\ref{sec:transit} I outline an example signal where transiting planets around many stars could act as a spatially distributed beacon network.
Finally in \S\ref{sec:discussion} I discuss ideas for other search approaches, and the advantages in cost and reproducibility of conducting SETI with current and upcoming optical time domain surveys.

@@

%%%%
\section{Optical Survey SETI}
\label{sec:method}

Wide-field imaging surveys, and their associated infrastructure, are an ideal data source for carrying out technosignature searches that augment efforts in other wavelengths or with different observing strategies. 
By covering large portions of the night sky with repeated imaging, and with surveys that span years to decades in duration, we can substantially reduce the challenge of deciding {\it when} and {\it where} to look. 
Increasingly, multiple imaging surveys obtain imaging that overlaps in both time and sky coverage, but with complimentary properties (e.g. imaging depth, cadence, wavelength coverage), which allows for rapid vetting and characterization of rare or low signal-to-noise events.. This also results in survey data providing some ``ground truth'' about the nature of most known objects in our Galaxy (e.g. their positions, temperatures, etc), and allows us to eliminate uncertainty on {\it what} we're searching for technosignatures.

% Reproducible science
The large archives of data produced from these surveys become valuable legacies that enable additional science long after data acquisition has concluded.

Studies for new phenomena in time-domain astronomy can therefore be carried out both in the future with new observations and in the past by re-analyzing survey archives.

%somrthing about sky surveys \citep{djorgovski2013}

%include mantras of "worth doing", "do it almost for free", "reproducible", "portable, do it everywhere", and "outliers are nature papers"

large ground-based surveys, such as LSST, are a highly ranked component of advancement in our current landscape something something
\citep{astro2010}.

real time tools being developed for LSST \citep{schwamb2019}

RNAAS about tools to model technosignatures from radio surveys
\citep{forgan2019}

\citep{lacki2019} looking for specular reflections of interplanetary objects, particularly in big new surveys
